\textsc{PAPERCLIP}: Associating Astronomical Observations and Natural Language with Multi-Modal Models

### 2024-03-11 paper/hubble_paperclip.tex
Finished related works

\def\preprintno{XXXX} % Insert correct preprint number

@@

% Define a new fancy page style
\fancypagestyle{firstpage}{
    \rhead{MIT-CTP/\preprintno}
}

@@

% \SM{Add some qualitative evals with the abstract model.}

@@

There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards}, with astrophysics being ripe for such an effort given the large amounts of publicly-available data and diverse modes of interacting with it.
%
The multi-modality inherent to astrophysical observations, with different types of data (e.g., images, spectra, light curves, textual descriptions) often available for a given target object, presents a unique opportunity.
%
This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} to construct a joint embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
%
\textsc{AstroLLaMA}~\citep{nguyen2023astrollama,perkowski2024astrollama} is another recent effort to fine-tune a publicly-available model (\textsc{Llama-2}) on astrophysics-specific textual data from the arXiv.

@@

\paragraph*{Related work}

% The CLIP family of foundation models, which in their original form embed images and associated captions into a common representation space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval~\citep{radford2021learning}.
%
The concept of learning task-agnostic representations via self-supervised learning has been applied within astrophysics \cite{slijepcevic2024radio,stein2021self,hayat2021self,slijepcevic2022learning} and used for downstream tasks like object similar search \citep{stein2021self}, gravitational lens finding \citep{stein2022mining}, estimation of Galactic distances \citep{hayat2021estimating}, and identification of rare galaxies \citep{walmsley2023rare}.
%
Associating different modalities via contrastive training has been employed in many other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations. Here, we present for the first time an application associating diverse astronomical data with the text modality.
%
For a recent review of contrastive learning in astrophysics, see \citet{huertas2023brief}. 
%
% \citet{bowles2023radio,bowles2022new}

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=0.99\textwidth]{plots/figure.pdf}
\caption{\changes{Overview of the PAPERCLIP method. (Left) A pre-trained CLIP model is fine-tuned using a dataset of \hubble observations and corresponding proposal abstracts. (Right) The fine-tuned model can then be used for downstream tasks such as observation retrieval (i.e., finding the observations most relevant to a given text query). The proposal abstract snippet here corresponds to proposal ID \href{https://archive.stsci.edu/proposal_search.php?id=16914&mission=hst}{16914}}.}
\label{fig:overview}
\end{figure*}

>>>>

@@

% Define a new fancy page style
% Insert correct preprint number
\def\preprintno{5690} 
\fancypagestyle{firstpage}{
    \rhead{MIT-CTP/\preprintno}
}

@@

@@

There is considerable interest in developing custom foundation models for the sciences \citep[e.g., ][]{batatia2023foundation,subramanian2023towards}, with astrophysics being ripe for such an effort given the large amounts of publicly-available data and diverse modes of interacting with it.
%
The multi-modality inherent to astrophysical observations, with different types of data (e.g., images, spectra, light curves, textual descriptions) often available for a given target object, presents a unique opportunity.
%
This multi-modality was recently exploited in \textsc{AstroCLIP}~\citep{lanusse2023astroclip} to construct a joint embedding space between multi-band images and optical spectra from the Dark Energy Spectroscopic Instrument (DESI).
%

@@

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=0.99\textwidth]{plots/figure.pdf}
\caption{\changes{Overview of the PAPERCLIP method. (Left) A pre-trained CLIP model is fine-tuned using a dataset of \hubble observations and corresponding proposal abstracts. The proposal abstracts are optionally summarized using guided large language model generation. (Right) The fine-tuned model can then be used for downstream tasks such as observation retrieval (i.e., finding the observations most relevant to a given text query). The proposal abstract snippet shown here corresponds to proposal ID \href{https://archive.stsci.edu/proposal_search.php?id=16914&mission=hst}{16914}}.}
\label{fig:overview}
\end{figure*}

\section{\changes{Related work}}

% The CLIP family of foundation models, which in their original form embed images and associated captions into a common representation space via contrastive learning, have shown strong performance and generalization capabilities on a variety of downstream tasks including zero-shot classification and image retrieval~\citep{radford2021learning}.
%
The concept of learning task-agnostic representations via self-supervised and contrastive learning has been applied within astrophysics \citep{slijepcevic2024radio,stein2021self,hayat2021self,slijepcevic2022learning} and used for downstream tasks like object similarity search \citep{stein2021self}, gravitational lens finding \citep{stein2022mining}, estimation of Galactic distances \citep{hayat2021estimating}, identification of rare galaxies \citep{walmsley2023rare}, and data compression \citep{akhmetzhanova2024data}. For a recent review of contrastive learning in astrophysics, see \citet{huertas2023brief}. 
%

Beyond applications to a single modality, \textsc{AstroCLIP}~\citep{lanusse2023astroclip} recently used contrastive learning to learn a joint representation between galaxy images and associated spectra, showing that the learned representation embodies relevant physical properties and can be effectively used for downstream tasks like redshift and mass estimation.
%

\citet{bowles2023radio,bowles2022new} introduced a method to associate radio galaxy images with a natural language description of their morphology by using human-generated descriptions, with the goal of deriving semantic morphology classes and using them for classification.

%
Associating diverse modalities via contrastive learning has been employed in many other scientific domains~\citep[e.g.,][]{liu2023text,Sanchez-Fernandez2022.11.17.516915,lanusse2023astroclip,cepeda2023geoclip}, and has been shown to be effective in learning semantically meaningful joint representations.

% While the interaction of astrophysical data with large language models is fairly nascent, \textsc{AstroLLaMA}~\citep{nguyen2023astrollama,perkowski2024astrollama} is a recent effort to fine-tune a publicly-available model (\textsc{Llama-2}) on astrophysics-specific textual data from the arXiv.

In this paper, we present for the first time an application associating target-agnostic astronomical data with the text modality, showing that this can be effectively accomplished through contrastive learning by leveraging observational proposals to inform text captions.
%
